---
title: "2 Training and Variable Selection"
author: "Paul Hively"
date: "April 14, 2016"
output: html_document
---

Source can be viewed [on GitHub](https://github.com/phively/gsb-sols-proj/blob/master/2%20Training%20and%20variable%20selection.Rmd)

# Logistic modeling

## First attempt

Begin by loading the data used for the visualizations in part 1.

```{r, message=F, warning=F, cache=F}
#### Run Rscript0 to load useful packages and functions ----
source("Rscript0 - libraries.R")
source("f.Wrangle.R")
```
```{r, message=F, warning=F, cache=F}
#### Model data cleanup ----
source("Rscript1a - modeling data appends.R")
```

* [Rscript0 - libraries.R](https://github.com/phively/gsb-sols-proj/blob/master/Rscript0%20-%20libraries.R)
* [f.Wrangle.R](https://github.com/phively/gsb-sols-proj/blob/master/f.Wrangle.R)
* [Rscript1a - modeling data appends.R](https://github.com/phively/gsb-sols-proj/blob/master/Rscript1a%20-%20modeling%20data%20appends.R)

Recall:

> The existing model ("JMod") has two components. First, the probability a solicitation $S_{i}$ comes in before the end of the fiscal year is a function of the current stage progress $c$ and time until the end of the year $t$, $P(S_{i}=1)=f(c_{i},t)$. The expected FRP in the current fiscal year is this probability times the expected amount, $E(FRP)=P(S_{i}=1)E(S_{i})$.
>
> Focus on the probability model. I assume that there should also be some effect due to the average close rate, season (e.g. people give more in December), planned ask amount, actual ask amount, and potentially interactions. Something like this:
>
> $$P(S_{ij}=1)=f(c_{i},t,\mu,t_{i}^{*},a_{j}^{*},a_{j})$$

Start by trying to directly predict the booked in FY probabilities. The simplest approach is to treat everything that isn't continuous as a factor and chuck it into a logistic regression model.

$$logit(p)=log\Big(\frac{p}{1-p}\Big)=\eta$$
$$E(\eta_{i})=c_{i}+t_{i}^{*}$$

In other words, the logit of the expected close rate is dependent on some intercept $c_{i}$ due to stage, and some month effect $t_{i}^{*}$.

```{r, message=F, warning=F, cache=F}
# Load some functions I wrote to calculate model diagnostics
source("f.Diagnostics.R")
```

* [f.Diagnostics.R](https://github.com/phively/gsb-sols-proj/blob/master/f.Diagnostics.R)

```{r, message=F, warning=F, cache=F}
## Model each stage by when it was asked 
glm.pln.1 <- glm(FY.Plan.Book ~ as.factor(month(mdat$Planning.Dt)), data=mdat, family=binomial())
glm.clr.1 <- glm(FY.Clear.Book ~ as.factor(month(mdat$Planning.Dt)), data=mdat, family=binomial())
glm.ask.1 <- glm(FY.Ask.Book ~ as.factor(month(mdat$Planning.Dt)), data=mdat, family=binomial())
glm.ora.1 <- glm(FY.Oral.Book ~ as.factor(month(mdat$Planning.Dt)), data=mdat, family=binomial())
modnames <- c("Plan","Clear","Ask","Oral")
# Compare each of the models
kable(CompareDeviances(models=list(glm.pln.1, glm.clr.1, glm.ask.1, glm.ora.1), modnames=modnames, debug=F), digits=3)
```

Month-by-month coefficients result in an unambiguously better model compared to just an intercept per stage.

JMod has the following structure. Cell entries are probability to close given that a solicitation is in the indicated stage and month, with an expected close date that fiscal year:

| Stage  | July-Feb| Mar-Apr |   May   |  June   |
|--------|:-------:|:-------:|:-------:|:-------:| 
| Plan   |  1/6    | 1/8     | 0       |   0     |
| Clear  |  1/3    | 1/6     | 1/8     |   0     |
| Ask    |  2/3    | 1/3     | 1/6     |   1/8   |
| Oral   |  1      | 1       | 1       |   1     |

Here are the confusion matrices, with a classification threshold of $p=.5$:

```{r, message=F, warning=F, cache=F}
# Count tables
ConfusionMatrix(list(glm.pln.1, glm.clr.1, glm.ask.1, glm.ora.1), modnames=modnames)
# Proportion tables
ConfusionMatrix(list(glm.pln.1, glm.clr.1, glm.ask.1, glm.ora.1), modnames=modnames, counts=F)
```

Compare errors for JMod versus the month intercepts logistic model by stage. My error function is the $L_{1}$ norm:

$$L_{1}(\epsilon)=\sum_{i=1}^{n}\epsilon_{i} = \sum_{i=1}^{n}\left|\hat{p}_{i}-p_{i}\right|$$

I also included the squared $L_{2}$ norm (equivalent to $RSS$ in OLS regression) for comparison, but it seems like an odd metric for range-restricted data.

```{r, message=F, warning=F, cache=F}
## Load functions for JMod
source("f.JMod.R")
```

* [f.JMod.R](https://github.com/phively/gsb-sols-proj/blob/master/f.JMod.R)

```{r, echo=F, message=F, warning=F, cache=F}
## Planning stage
# Residuals
jmod.1 <- JMod(curr.stage=1, curr.dt=mdat$Planning.Dt, expected.dt=mdat$Expected.Dt)
jmod <- na.omit(JModError(probs=jmod.1$prob, truth=mdat$FY.Plan.Book))
glmod <- JModError(probs=glm.pln.1$fitted, truth=model.frame(glm.pln.1)[,1])
# Combined residuals
ggdat <- data.frame(jmod, glm=glmod, closed=na.omit(mdat$FY.Plan.Book))
ggplot(ggdat, aes(x=glm, y=jmod, color=closed)) + geom_point() + geom_hline(yintercept=mean(jmod), alpha=.3) + geom_vline(xintercept=mean(glmod), alpha=.3) + labs(title="Plan JMod versus glm residual error and bias") + xlim(c(-1,1)) + ylim(c(-1,1))
# Metrics
matrix(c(LpNorm(jmod, p=1), LpNorm(jmod)^2, LpNorm(glmod, p=1), LpNorm(glmod)^2), nrow=2, dimnames=list(c("L1","L2"),c("JMod","glm")))
```

JMod wins on the sum of absolute residuals, but it is very negatively biased, to the tune of $-0.325$.

```{r, echo=F, message=F, warning=F, cache=F}
## Clear stage
# Residuals
jmod.2 <- JMod(curr.stage=2, curr.dt=mdat$Clear.Dt, expected.dt=mdat$Expected.Dt)
jmod <- na.omit(JModError(probs=jmod.2$prob, truth=mdat$FY.Clear.Book))
glmod <- JModError(probs=glm.clr.1$fitted, truth=model.frame(glm.clr.1)[,1])
# Combined residuals
ggdat <- data.frame(jmod, glm=glmod, closed=na.omit(mdat$FY.Clear.Book))
ggplot(ggdat, aes(x=glm, y=jmod, color=closed)) + geom_point() + geom_hline(yintercept=mean(jmod), alpha=.3) + geom_vline(xintercept=mean(glmod), alpha=.3) + labs(title="Clear JMod versus glm residual error and bias") + xlim(c(-1,1)) + ylim(c(-1,1))
# Metrics
matrix(c(LpNorm(jmod, p=1), LpNorm(jmod)^2, LpNorm(glmod, p=1), LpNorm(glmod)^2), nrow=2, dimnames=list(c("L1","L2"),c("JMod","glm")))
```

A tiny bit closer on the sum of absolute residuals, but JMod is still negatively biased, around $-0.295$

```{r, echo=F, message=F, warning=F, cache=F}
## Ask stage
# Residuals
jmod.3 <- JMod(curr.stage=3, curr.dt=mdat$Ask.Dt, expected.dt=mdat$Expected.Dt)
jmod <- na.omit(JModError(probs=jmod.3$prob, truth=mdat$FY.Ask.Book))
glmod <- JModError(probs=glm.ask.1$fitted, truth=model.frame(glm.ask.1)[,1])
# Combined residuals
ggdat <- data.frame(jmod, glm=glmod, closed=na.omit(mdat$FY.Ask.Book))
ggplot(ggdat, aes(x=glm, y=jmod, color=closed)) + geom_point() + geom_hline(yintercept=mean(jmod), alpha=.3) + geom_vline(xintercept=mean(glmod), alpha=.3) + labs(title="Ask JMod versus glm residual error and bias") + xlim(c(-1,1)) + ylim(c(-1,1))
# Metrics
matrix(c(LpNorm(jmod, p=1), LpNorm(jmod)^2, LpNorm(glmod, p=1), LpNorm(glmod)^2), nrow=2, dimnames=list(c("L1","L2"),c("JMod","glm")))
```

Clear win for JMod; bias is only $-0.104$.

```{r, echo=F, message=F, warning=F, cache=F}
## Oral stage
# Residuals
jmod.4 <- JMod(curr.stage=4, curr.dt=mdat$Oral.Dt, expected.dt=mdat$Expected.Dt)
jmod <- na.omit(JModError(probs=jmod.4$prob, truth=mdat$FY.Oral.Book))
glmod <- JModError(probs=glm.ora.1$fitted, truth=model.frame(glm.ora.1)[,1])
# Combined residuals
ggdat <- data.frame(jmod, glm=glmod, closed=na.omit(mdat$FY.Oral.Book))
ggplot(ggdat, aes(x=glm, y=jmod, color=closed)) + geom_point() + geom_hline(yintercept=mean(jmod), alpha=.3) + geom_vline(xintercept=mean(glmod), alpha=.3) + labs(title="Oral JMod versus glm residual error and bias") + xlim(c(-1,1)) + ylim(c(-1,1))
# Metrics
matrix(c(LpNorm(jmod, p=1), LpNorm(jmod)^2, LpNorm(glmod, p=1), LpNorm(glmod)^2), nrow=2, dimnames=list(c("L1","L2"),c("JMod","glm")))
```

JMod bias is $0.047$.

Basically, what this comes down to is how concerned are we about bias? The negative bias is due to overaggressive discounting as the end of the fiscal year (June 30) approaches.

One more idea for model comparison: why not look at the classification tables directly? Since JMod has hard cutoffs it doesn't really make sense to use a threshold. Instead, add the predicted probabilities to the corresponding cells of a confusion matrix for each observation.

```{r, message=F, warning=F, cache=F}
## Confusion matrix summing probabilities, rather than counting classifications
# GLM for plan stage
ConfusionMatrixWeighted(preds=fitted(glm.pln.1), truth=model.frame(glm.pln.1)[,1], dimnames=c("truth","pred"))
# JMod for plan stage
(jmod.1$errtable <- ConfusionMatrixWeighted(preds=jmod.1$prob, truth=mdat$FY.Plan.Book, dimnames=c("truth","pred")))
```

Note that the antidiagonal terms sum to the corresponding $L_{1}$ norm, from the "Plan JMod versus glm residual error and bias" table above.

Planned solicitations -- while the glm is well balanced, JMod overwhelmingly classifies planned solicitations as unlikely to come in. This could be because *JMod is tuned toward directly predicting ask amounts*, and uses its weights to discount MG rather than worrying about the close probability of every AF gift.

```{r, message=F, warning=F, cache=F}
# GLM for clear stage
ConfusionMatrixWeighted(preds=fitted(glm.clr.1), truth=model.frame(glm.clr.1)[,1], dimnames=c("truth","pred"))
# JMod for clear stage
(jmod.2$errtable <- ConfusionMatrixWeighted(preds=na.omit(jmod.2$prob), truth=na.omit(mdat$FY.Clear.Book), dimnames=c("truth","pred")))
```

Same story -- for cleared solicitations JMod overwhelmingly categorizes them as unlikely to close.

```{r, message=F, warning=F, cache=F}
# GLM for ask stage
ConfusionMatrixWeighted(preds=fitted(glm.ask.1), truth=model.frame(glm.ask.1)[,1], dimnames=c("truth","pred"))
# JMod for ask stage
(jmod.3$errtable <- ConfusionMatrixWeighted(preds=na.omit(jmod.3$prob), truth=na.omit(mdat$FY.Ask.Book), dimnames=c("truth","pred")))
```

Same story for asked, but look how much better JMod does in the True-True quadrant, bottom right.

```{r, message=F, warning=F, cache=F}
# GLM for oral stage
ConfusionMatrixWeighted(preds=fitted(glm.ora.1), truth=model.frame(glm.ora.1)[,1], dimnames=c("truth","pred"))
# JMod for oral stage
(jmod.4$errtable <- ConfusionMatrixWeighted(preds=na.omit(jmod.4$prob), truth=na.omit(mdat$FY.Oral.Book), dimnames=c("truth","pred")))
```

Oral is different; JMod does better in every quadrant. In particular, False-False in the top left is about twice as good as glm. Under JMod, an oral pledge is only ever assigned $p=0$ when it is scheduled to be closed in a future fiscal year.

Conclusion: *the most important* variable not currently captured by glm is whether the expected date is in a future year.

# Variable selection

Definitely include an indicator for **expected date in the future**.

From [part 1](https://github.com/phively/gsb-sols-proj/blob/master/1%20Data%20exploration%20commentary.Rmd), some interesting variables include **expected amount**, and whether **ask amount/planned ask** is greater or less than 1.

Potentially interesting derived variables include the **number of days left in the year**, the **number of days until expected close**, and the **amount of time that's passed since the previous stage**.

Some other thoughts.

1. Is logistic regression really suitable (overdispersion)?
2. Should it be restricted to MG-range gifts, $[\$25\text{k},\$5\text{M})$?
3. What nonparametric methods are worth trying?

First, let's create those derived variables.

```{r, message=F, warning=F, cache=F}
#### Derived variables ----
source("Rscript1b - derived variables.R")
```

* [Rscript1b - derived variables.R](https://github.com/phively/gsb-sols-proj/blob/master/Rscript1b%20-%20derived%20variables.R)

Think about #3 first. I'm a fan of classification trees, and that general method is the closest to the existing approach. It would also be interesting to see which of the proposed variables end up being chosen.

Try the random forest approach to select variables. First split up the variables by which stages it's suitable to use them for prediction.

```{r, message=F, warning=F, cache=F}
## Identify which dependent variables contain information specific to the various stages
## Obviously shouldn't include after-the-fact information (FY. objects, actual amount, etc.)
head.all <- names(mderived)[names(mderived) %in% c("Solicitation.Type.Desc", "Planned.Fiscal.Mo", "Planned.Amt", "Expected.Fiscal.Mo", "Expected.Amt", "Planned.Ask.Band")]
# Look for "plan" and "Plan" in header ([pP] = both cases), but ignore the FY. outcome objects
head.plan <- names(mderived)[setdiff(grep("[pP]lan", names(mderived)), grep("FY.", names(mderived)))]
head.plan <- setdiff(head.plan, c(head.all, "plan2clear")) #take plan2clear which wouldn't have been known
# Look for "Clear" in header
head.clear <- names(mderived)[setdiff(grep("[cC]lear", names(mderived)), grep("FY.", names(mderived)))]
head.clear <- setdiff(head.clear, c(head.all, "clear2ask")) #take out clear2ask which wouldn't have been known
# Look for "Ask" in header
head.ask <- names(mderived)[setdiff(grep("[aA]sk", names(mderived)), grep("FY.", names(mderived)))]
head.ask <- setdiff(head.ask, c(head.all, "ask2oral")) #take out ask2oral which wouldn't have been known
head.ask <- c(head.ask, "plan2clear") #add fields that would have been known
# Look for "Oral" in header
head.oral <- names(mderived)[setdiff(grep("[oO]ral", names(mderived)), grep("FY.", names(mderived)))]
head.oral <- setdiff(head.oral, head.all)
head.oral <- c(head.oral, "plan2clear", "clear2ask", "Ask.Amt", "Ask.Band", "Ask.Amt.Over.Pln.Amt", "Ask.Amt.Over.Pln.Fac") #add fields that would've been known
```

Now grow a random forest for each stage.

```{r, message=F, warning=F, cache=F}
library(randomForest)
set.seed(15584)
# Planned stage forest
treedat <- na.omit(mderived[,c("FY.Plan.Book", head.all, head.plan)])
rf.plan <- randomForest(factor(FY.Plan.Book) ~ ., data=treedat, importance=T)
# Clear stage forest
treedat <- na.omit(mderived[,c("FY.Clear.Book", head.all, head.clear)])
rf.clear <- randomForest(factor(FY.Clear.Book) ~ ., data=treedat, importance=T)
# Ask stage forest
treedat <- na.omit(mderived[,c("FY.Ask.Book", head.all, head.ask)])
rf.ask <- randomForest(factor(FY.Ask.Book) ~ ., data=treedat, importance=T)
# Oral stage forest
treedat <- na.omit(mderived[,c("FY.Oral.Book", head.all, head.oral)])
rf.oral <- randomForest(factor(FY.Oral.Book) ~ ., data=treedat, importance=T)
```

The green lines are the out-of-bag error (similar to cross-validation). For comparison, I've included the error rates for the initial logistic regression, above. (Note that `PlotForest()` is included in [f.Diagnostics.R](https://github.com/phively/gsb-sols-proj/blob/master/f.Diagnostics.R).)

```{r, message=F, warning=F, cache=F}
PlotForest(rf.plan, title="Plan")
PlotForest(rf.clear, title="Clear")
PlotForest(rf.ask, title="Ask")
PlotForest(rf.oral, title="Oral")
```

Finally, the GLM errors (from the confusion matrix, threshold $p=0.5$) versus the JMod errors (from the weighted confusion matrix, $1-\sum_{i=1}^{n}\frac{\mathbf{1}\{x_{i}\text{ misclassified}\}}{n}$) versus the out-of-bag decision forest errors:

```{r, echo=F, message=F, warning=F, cache=F}
# Put together the confusion matrices again
confuse <- ConfusionMatrix(list(glm.pln.1, glm.clr.1, glm.ask.1, glm.ora.1), modnames=modnames, counts=F)
# First logistic regression errors
dat1 <- data.frame(Model="GLM", Plan=confuse[["Plan"]]["FALSE","TRUE"] + confuse[["Plan"]]["TRUE","FALSE"], Clear=confuse[["Clear"]]["FALSE","TRUE"] + confuse[["Clear"]]["TRUE","FALSE"], Ask=confuse[["Ask"]]["FALSE","TRUE"] + confuse[["Ask"]]["TRUE","FALSE"], Oral=tryCatch(confuse[["Oral"]]["FALSE","TRUE"] + confuse[["Oral"]]["TRUE","FALSE"], error=function(.) sum(confuse[["Oral"]]["FALSE","TRUE"])))
## JMod errors
dat2 <- data.frame(Model="JMod", Plan=1-sum(diag(jmod.1$errtable))/sum(jmod.1$errtable), Clear=1-sum(diag(jmod.2$errtable))/sum(jmod.2$errtable), Ask=1-sum(diag(jmod.3$errtable))/sum(jmod.3$errtable), Oral=1-sum(diag(jmod.4$errtable))/sum(jmod.4$errtable))
## randomForest errors
dat3 <- data.frame(Model="Random forest", Plan=tail(rf.plan$err.rate[,"OOB"],1), Clear=tail(rf.clear$err.rate[,"OOB"],1), Ask=tail(rf.ask$err.rate[,"OOB"],1), Oral=tail(rf.oral$err.rate[,"OOB"],1))
## error table
errors <- rbind(dat1, dat2, dat3)
kable(errors, digits=2)
```

Note the great out-of-bag error rates across the board. The random forests are definitely doing something right. Which variables do they find the most helpful?

```{r, echo=F, message=F, warning=F, cache=F}
## Coerce importances into a data frame
imp.plan <- importance(rf.plan, scale=F, type=1)
imp.clear <- importance(rf.clear, scale=F, type=1)
imp.ask <- importance(rf.ask, scale=F, type=1)
imp.oral <- importance(rf.oral, scale=F, type=1)
rf.importances <- rbind(data.frame(Model="Plan", imp.plan, Variable=rownames(imp.plan)), data.frame(Model="Clear", imp.clear, Variable=rownames(imp.clear)), data.frame(Model="Ask", imp.ask, Variable=rownames(imp.ask)), data.frame(Model="Oral", imp.oral, Variable=rownames(imp.oral)), make.row.names=F)
## Plot importances
# y axis will end at nearest .02 above the maximum in the data
top <- ceiling(max(rf.importances$MeanDecreaseAccuracy)*50)/50
ggplot(rf.importances, aes(x=Variable, y=MeanDecreaseAccuracy, color=Model, group=Model)) + geom_point() + geom_line() + theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(title="Variable selection", y="Mean Decrease in Accuracy") + scale_y_continuous(breaks=seq(0, top, by=.02))
```

* The `Stage.Future`, `stage2EOFY`, `stage.fiscal.mo`, and `stage2expect` variables are correlated. Because of how the forests are constructed (random branching) that means they all show up as fairly important.
* Out of the four of them I would personally ditch `stage2EOFY` and consider dropping one of the `stage.fiscal.mo` or `stage2expect` variables as well.
* `Expected.Amt` and `Planned.Amt` are good predictors before the Oral stage; `Planned.Ask.Band` is worse across the board.
* `Expected.Fiscal.Mo` is important across the board, while `Planned.Fiscal.Mo` is useful before the Oral stage.
* Similarly, `Solicitation.Type.Desc` is a decent predictor before 
* The various times between stages e.g. `plan2clear` are not good predictors.
* `Ask.Amt` is a good predictor in the Ask stage, certainly much better than `Ask.Band`; the two `Ask.Amt.Over.` fractions are useless.

```{r, message=F, warning=F, cache=F}
## Update the explanatory variables
exclude <- c("plan2EOFY", "clear2EOFY", "ask2EOFY", "oral2EOFY", "plan2clear", "clear2ask", "ask2oral", "Planned.Ask.Band", "Ask.Band", "Ask.Amt.Over.Pln.Amt", "Ask.Amt.Over.Pln.Fac")
head.all <- setdiff(head.all, exclude)
head.plan <- setdiff(head.plan, exclude)
head.clear <- setdiff(head.clear, exclude)
head.ask <- setdiff(head.ask, exclude)
head.oral <- setdiff(head.oral, exclude)
## Regrow the random forests
set.seed(66461)
# Planned stage forest
treedat <- na.omit(mderived[,c("FY.Plan.Book", head.all, head.plan)])
rf.plan <- randomForest(factor(FY.Plan.Book) ~ ., data=treedat, importance=T)
# Clear stage forest
treedat <- na.omit(mderived[,c("FY.Clear.Book", head.all, head.clear)])
rf.clear <- randomForest(factor(FY.Clear.Book) ~ ., data=treedat, importance=T)
# Ask stage forest
treedat <- na.omit(mderived[,c("FY.Ask.Book", head.all, head.ask)])
rf.ask <- randomForest(factor(FY.Ask.Book) ~ ., data=treedat, importance=T)
# Oral stage forest
treedat <- na.omit(mderived[,c("FY.Oral.Book", head.all, head.oral)])
rf.oral <- randomForest(factor(FY.Oral.Book) ~ ., data=treedat, importance=T)
```

Plotting the new importances:

```{r, echo=F, message=F, warning=F, cache=F}
## Coerce importances into a data frame
imp.plan <- importance(rf.plan, scale=F, type=1)
imp.clear <- importance(rf.clear, scale=F, type=1)
imp.ask <- importance(rf.ask, scale=F, type=1)
imp.oral <- importance(rf.oral, scale=F, type=1)
rf.importances <- rbind(data.frame(Model="Plan", imp.plan, Variable=rownames(imp.plan)), data.frame(Model="Clear", imp.clear, Variable=rownames(imp.clear)), data.frame(Model="Ask", imp.ask, Variable=rownames(imp.ask)), data.frame(Model="Oral", imp.oral, Variable=rownames(imp.oral)), make.row.names=F)
## Plot importances
# y axis will end at nearest .02 above the maximum in the data
top <- ceiling(max(rf.importances$MeanDecreaseAccuracy)*50)/50
ggplot(rf.importances, aes(x=Variable, y=MeanDecreaseAccuracy, color=Model, group=Model)) + geom_point() + geom_line() + theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(title="Variable selection", y="Mean Decrease in Accuracy") + scale_y_continuous(breaks=seq(0, top, by=.02))
```

And the new errors table:

```{r, echo=F, message=F, warning=F, cache=F}
## randomForest errors
errors <- rbind(errors, data.frame(Model="Random forest 2", Plan=tail(rf.plan$err.rate[,"OOB"],1), Clear=tail(rf.clear$err.rate[,"OOB"],1), Ask=tail(rf.ask$err.rate[,"OOB"],1), Oral=tail(rf.oral$err.rate[,"OOB"],1), stringsAsFactors=F))
## error table
kable(errors, digits=2)
```

Looks like I might even be able to get away with taking out some more predictors. What if we exclude the redundant `stage2expect`, and `Expected.Fiscal.Mo` and `Planned.Fiscal.Mo` which are overwhelmingly December or June?

```{r, message=F, warning=F, cache=F}
## Update the explanatory variables
exclude <- c(exclude, "plan2expect", "clear2expect", "ask2expect", "oral2expect", "Expected.Fiscal.Mo", "Planned.Fiscal.Mo")
head.all <- setdiff(head.all, exclude)
head.plan <- setdiff(head.plan, exclude)
head.clear <- setdiff(head.clear, exclude)
head.ask <- setdiff(head.ask, exclude)
head.oral <- setdiff(head.oral, exclude)
## Regrow the random forests
set.seed(48676)
# Planned stage forest
treedat <- na.omit(mderived[,c("FY.Plan.Book", head.all, head.plan)])
rf.plan <- randomForest(factor(FY.Plan.Book) ~ ., data=treedat, importance=T)
# Clear stage forest
treedat <- na.omit(mderived[,c("FY.Clear.Book", head.all, head.clear)])
rf.clear <- randomForest(factor(FY.Clear.Book) ~ ., data=treedat, importance=T)
# Ask stage forest
treedat <- na.omit(mderived[,c("FY.Ask.Book", head.all, head.ask)])
rf.ask <- randomForest(factor(FY.Ask.Book) ~ ., data=treedat, importance=T)
# Oral stage forest
treedat <- na.omit(mderived[,c("FY.Oral.Book", head.all, head.oral)])
rf.oral <- randomForest(factor(FY.Oral.Book) ~ ., data=treedat, importance=T)
## randomForest errors
errors <- rbind(errors, data.frame(Model="Random forest 3", Plan=tail(rf.plan$err.rate[,"OOB"],1), Clear=tail(rf.clear$err.rate[,"OOB"],1), Ask=tail(rf.ask$err.rate[,"OOB"],1), Oral=tail(rf.oral$err.rate[,"OOB"],1), stringsAsFactors=F))
## error table
kable(errors, digits=2)
```

Pretty much the same as the other forest methods, so I prefer the more parsimonious model. Finally, how do random forests compare to JMod when given the same amount of information?

```{r, message=F, warning=F, cache=F}
## Regrow the random forests
set.seed(30635)
# Planned stage forest
treedat <- na.omit(mderived[,c("FY.Plan.Book", "planning.fiscal.mo", "plan2EOFY")])
rf.plan <- randomForest(factor(FY.Plan.Book) ~ ., data=treedat, importance=T)
# Clear stage forest
treedat <- na.omit(mderived[,c("FY.Clear.Book", "clear.fiscal.mo", "clear2EOFY")])
rf.clear <- randomForest(factor(FY.Clear.Book) ~ ., data=treedat, importance=T)
# Ask stage forest
treedat <- na.omit(mderived[,c("FY.Ask.Book", "ask.fiscal.mo", "ask2EOFY")])
rf.ask <- randomForest(factor(FY.Ask.Book) ~ ., data=treedat, importance=T)
# Oral stage forest
treedat <- na.omit(mderived[,c("FY.Oral.Book", "oral.fiscal.mo", "oral2EOFY")])
rf.oral <- randomForest(factor(FY.Oral.Book) ~ ., data=treedat, importance=T)
## randomForest errors
errors <- rbind(errors, data.frame(Model="Random forest, JMod fields only", Plan=tail(rf.plan$err.rate[,"OOB"],1), Clear=tail(rf.clear$err.rate[,"OOB"],1), Ask=tail(rf.ask$err.rate[,"OOB"],1), Oral=tail(rf.oral$err.rate[,"OOB"],1), stringsAsFactors=F))
## error table
kable(errors, digits=2)
```

Interesting -- limiting random forest to JMod fields (allows splits only on `Stage`, `stage.fiscal.mo`, and `stage2EOFY`) results in a model that doesn't do any better than the basic GLM. Random forest 3 is parsimonious so it looks like we may have a final set of variables.

## Plan

```{r, echo=F, message=F, warning=F, cache=F}
c(head.all, head.plan)
```

## Clear

```{r, echo=F, message=F, warning=F, cache=F}
c(head.all, head.clear)
```

## Ask

```{r, echo=F, message=F, warning=F, cache=F}
c(head.all, head.ask)
```

## Oral

```{r, echo=F, message=F, warning=F, cache=F}
c(head.all, head.oral)
```

# Packages used
```{r}
session_info()
```