---
title: "4 Head to head models"
author: "Paul Hively"
date: "June 3, 2016"
output: html_document
---

Source can be viewed [on GitHub](https://github.com/phively/gsb-sols-proj/blob/master/4%20Head%20to%20head%20models.Rmd)

# Loading and formatting the data

Begin by loading the models to be compared.

```{r, message=F, warning=F, cache=F}
#### Run Rscript0 to load useful packages and functions ----
source("Rscript0 - libraries.R")
source("f.Wrangle.R")
source("f.Diagnostics.R")
source("f.JMod.R")
```
```{r, message=F, warning=F, cache=F}
#### Model data cleanup ----
source("Rscript1a - modeling data appends.R")
source("Rscript1b - derived variables.R")
source("Rscript1c - additional predictors.R")
#### Models from last time ----
load("logistic.models.Rdata")
### Point-in-time data file ----
sols <- read.csv("sols.dedupe.by.mo.txt", sep="\t", stringsAsFactors=F)
```

* [Rscript0 - libraries.R](https://github.com/phively/gsb-sols-proj/blob/master/Rscript0%20-%20libraries.R)
* [f.Wrangle.R](https://github.com/phively/gsb-sols-proj/blob/master/f.Wrangle.R)
* [f.Diagnostics.R](https://github.com/phively/gsb-sols-proj/blob/master/f.Diagnostics.R)
* [f.JMod.R](https://github.com/phively/gsb-sols-proj/blob/master/f.JMod.R)
* [Rscript1c - additional predictors.R](https://github.com/phively/gsb-sols-proj/blob/master/Rscript1c%20-%20additional%2predictors.R)

# Data transformations and setup for the model

The `sols` data frame was created for a different project and doesn't have all the columns needed to make predictions in the new model. The following fields are required:

#### JMod
`Curr.Stage`, `Curr.Dt`, `Expected.Dt`, `Expected.Amt`, `Actual.Amt`, `Closed.In.FY`

#### GLMs
`FY.Plan.Book`, `FY.Clear.Book`, `FY.Ask.Book`, `FY.Oral.Book`, `Sol.Type.Agg`, `lt.Planned.Amt`, `lt.Expected.Amt`, `lt.Ask.Amt`, `Plan.Future`, `Clear.Future`, `Ask.Future`, `Oral.Future`

Let's go about reshaping `sols`:

```{r, message=F, warning=F, cache=F}
## Grab existing fields
sols2 <- sols %>% select(Solicitation.ID, Final.Sol.Stage, Solicitation.Stage.Desc, filename, rpt.date, Expected.Dt, Expected.Amt, Actual.Amt, Resolved.In.FY, Closed.In.FY, Solicitation.Type.Desc, Planned.Ask.Amt, Ask.Amt)
## Convert date fields
sols2 <- ToDate(sols2, c("rpt.date", "Expected.Dt"))
## Impute any missing values from mdat
# Expected
sols2$Expected.Amt[is.na(sols2$Expected.Amt)] <- left_join(sols2[is.na(sols2$Expected.Amt), c("Solicitation.ID", "Expected.Amt")], mdat[, c("Solicitation.ID", "Expected.Amt")], by="Solicitation.ID")[, 3]
# Actual
sols2$Actual.Amt[is.na(sols2$Actual.Amt)] <- left_join(sols2[is.na(sols2$Actual.Amt), c("Solicitation.ID", "Actual.Amt")], mdat[, c("Solicitation.ID", "Actual.Amt")], by="Solicitation.ID")[, 3]
# Planned
sols2$Planned.Ask.Amt[is.na(sols2$Planned.Ask.Amt)] <- left_join(sols2[is.na(sols2$Planned.Ask.Amt), c("Solicitation.ID", "Planned.Ask.Amt")], mdat[, c("Solicitation.ID", "Planned.Amt")], by="Solicitation.ID")[, 3]
# Ask
sols2$Ask.Amt[is.na(sols2$Ask.Amt)] <- left_join(sols2[is.na(sols2$Ask.Amt), c("Solicitation.ID", "Ask.Amt")], mdat[, c("Solicitation.ID", "Ask.Amt")], by="Solicitation.ID")[, 3]
## Create derived fields
sol.stage <- substr(sols2$Solicitation.Stage.Desc, 1, 2)
sols2 <- sols2 %>% mutate(
  Curr.Stage = ifelse(sol.stage %in% c("1.", "2.", "3.", "3A", "3B", "3C", "4."), 1,
               ifelse(sol.stage %in% c("5.", "6."), 2,
               ifelse(sol.stage %in% c("7."), 3,
               ifelse(sol.stage %in% c("8."), 4,
               ifelse(sol.stage %in% c("9."), 5,
               NA)
               )))),
  FY.Plan.Book = Curr.Stage == 1 & Closed.In.FY,
  FY.Clear.Book = Curr.Stage == 2 & Closed.In.FY,
  FY.Ask.Book = Curr.Stage == 3 & Closed.In.FY,
  FY.Oral.Book = Curr.Stage == 4 & Closed.In.FY,
  Sol.Type.Agg = factor(
	  ifelse(Solicitation.Type.Desc == "Standard Pledge", "Standard Pledge",
	  ifelse(Solicitation.Type.Desc == "Outright Gift", "Outright Gift",
	  "Other")
  )),
  lt.Planned.Amt = LogTrans(Planned.Ask.Amt),
  lt.Expected.Amt = LogTrans(Expected.Amt),
  lt.Ask.Amt = LogTrans(Ask.Amt),
  Actual.Amt.In.FY = Actual.Amt * Closed.In.FY,
  Expect.In.Future.FY = Expected.Dt > FYE(rpt.date)
)
## Select needed fields
sols2 <- sols2 %>% select(Solicitation.ID, Final.Sol.Stage, Solicitation.Stage.Desc, Curr.Stage, filename, rpt.date, Expected.Dt, Planned.Ask.Amt, Ask.Amt, Expected.Amt, Actual.Amt, Actual.Amt.In.FY, Closed.In.FY, Sol.Type.Agg, lt.Planned.Amt, lt.Expected.Amt, lt.Ask.Amt, Expected.Dt, Expect.In.Future.FY)
write.table(sols2, file="pit-dat.txt", sep="\t", row.names=F)
```

# Comparisons

Now that I have a point-in-time data file, I can make predictions for each discrete time point.

```{r, message=F, warning=F, cache=F}
sols3 <- na.omit(sols2 %>% select(Solicitation.ID, Curr.Stage, rpt.date, Expected.Dt, Expected.Amt, Actual.Amt, Closed.In.FY, Expect.In.Future.FY, Sol.Type.Agg, lt.Planned.Amt, lt.Expected.Amt, lt.Ask.Amt))
## JMod predictions
jmod.preds <- JMod(curr.stage=sols3$Curr.Stage, curr.dt=sols3$rpt.date, expected.dt=sols3$Expected.Dt, expect.amt=sols3$Expected.Amt, act.amt=sols3$Actual.Amt, closed.in.fy=sols3$Closed.In.FY)
# Add the report date
jmod.preds$dt <- sols3$rpt.date
jmod.preds$id <- as.character(paste(sols3$Solicitation.ID, sols3$rpt.date, sep="."))
jmod.preds$expect <- sols3$Expected.Amt
jmod.preds <- data.frame(jmod.preds)
## GLM predictions
PredictGLM <- function(model, data){
  data.frame(probability = predict(model, newdata=data, type="response")) %>%
  mutate(prediction = probability * data$Expected.Amt,
         error = prediction - ifelse(is.na(data$Actual.Amt), 0, data$Actual.Amt) * data$Closed.In.FY,
         dt = data$rpt.date,
         expect = data$Expected.Amt)
}
# Plan
glmdat <- sols3 %>% filter(Curr.Stage == 1) %>% mutate(FY.Plan.Book = Closed.In.FY, Plan.Future = Expect.In.Future.FY)
plan.preds <- PredictGLM(glm.pln.t, glmdat)
# Clear
glmdat <- sols3 %>% filter(Curr.Stage == 2) %>% mutate(FY.Clear.Book = Closed.In.FY, Clear.Future = Expect.In.Future.FY)
clear.preds <- PredictGLM(glm.clr.t, glmdat)
# Ask
glmdat <- sols3 %>% filter(Curr.Stage == 3) %>% mutate(FY.Ask.Book = Closed.In.FY, Ask.Future = Expect.In.Future.FY)
ask.preds <- PredictGLM(glm.ask.t, glmdat)
# Oral
glmdat <- sols3 %>% filter(Curr.Stage == 4) %>% mutate(FY.Oral.Book = Closed.In.FY, Oral.Future = Expect.In.Future.FY)
oral.preds <- PredictGLM(glm.ora.t, glmdat)
# Combine
glm.preds <- rbind(plan.preds, clear.preds, ask.preds, oral.preds)
```

Take a look at the comparison.

```{r, echo=F, message=F, warning=F, cache=F}
## Set up comparison data frame
ggdat <- jmod.preds %>% mutate(dt = DateToNthOfMonth(dt, n=1)) %>% group_by(dt) %>% summarise(pred = sum(prediction), act = sum(actual))
## Reformat glm predictions
tmpdf <- glm.preds %>% mutate(dt = DateToNthOfMonth(dt, n=1)) %>% group_by(dt) %>% summarise(glm.pred = sum(prediction))
## Bind tmpdf to ggdat
ggdat <- left_join(ggdat, tmpdf, by = "dt")
## Plot the comparison
ggplot(data=ggdat, aes(x=dt, y=act)) + geom_point(aes(color="Actual")) + geom_line(aes(color="Actual"), size=1.5, alpha=.5) + geom_point(aes(y=pred, color="JMod")) + geom_line(aes(y=pred, color="JMod")) + geom_point(aes(y=glm.pred, color="GLM")) + geom_line(aes(y=glm.pred, color="GLM")) + scale_y_continuous(name="Amount", labels=scales::dollar) + scale_x_date(name="Date", date_minor_breaks="month") + labs(title="All solicitations")
```

This includes PG solicitations; the GLM way overshoots in FY15 on.

```{r, echo=F, message=F, warning=F, cache=F}
## Set up comparison data frame
x <- 5000000
ggdat <- filter(jmod.preds, expect < x) %>% mutate(dt = DateToNthOfMonth(dt, n=1)) %>% group_by(dt) %>% summarise(pred = sum(prediction), act = sum(actual))
## Reformat glm predictions
tmpdf <- filter(glm.preds, expect < x) %>% mutate(dt = DateToNthOfMonth(dt, n=1)) %>% group_by(dt) %>% summarise(glm.pred = sum(prediction))
## Bind tmpdf to ggdat
ggdat <- left_join(ggdat, tmpdf, by = "dt")
## Plot the comparison
ggplot(data=ggdat, aes(x=dt, y=act)) + geom_point(aes(color="Actual")) + geom_line(aes(color="Actual"), size=1.5, alpha=.5) + geom_point(aes(y=pred, color="JMod")) + geom_line(aes(y=pred, color="JMod")) + geom_point(aes(y=glm.pred, color="GLM")) + geom_line(aes(y=glm.pred, color="GLM")) + scale_y_continuous(name="Amount", labels=scales::dollar) + scale_x_date(name="Date", date_minor_breaks="month") + labs(title=paste("Solicitations under", dollar(x)))
```

The GLM is clearly better pre-FY15 and after that it seems like a toss-up. That makes sense considering the JMod coefficients are revised each fiscal year while the GLM was trained on the whole dataset.

Of course, the as-yet-unmentioned advantage of using a probabalistic approach is that it allows for the easy estimation of confidence intervals.

```{r, message=F, warning=F, cache=F}
## Bootstrapped GLM dataset
set.seed(2358)
rep <- 1000
boot.result <- matrix(0, nrow = nrow(glm.preds), ncol = rep)
for (i in 1:rep) {
  boot.result[, i] <- (glm.preds$probability >= runif(n=nrow(glm.preds), min=0, max=1)) * glm.preds$expect
}
glm.boot <- cbind(glm.preds, data.frame(boot.result))
# Summarise by date across every bootstrapped dataset
glm.boot <- filter(glm.boot, expect < x) %>% mutate(dt = DateToNthOfMonth(dt, n=1)) %>% group_by(dt) %>% summarise_each(funs(sum))
# Find empirical 2.5th, 50th, and 97.5th percentiles (95% CI)
boot.names <- paste("X", 1:rep, sep="")
ggdat <- cbind(ggdat, t(apply(glm.boot[, boot.names], MARGIN=1, FUN=quantile, probs=c(.025, .5, .975))))
```

Basically, for each solicitation, draw `rep` samples from the $\text{Unif}(0,1)$ distribution, and if the randomly generated number is less than or equal to the predicted probability the solicitation will close, count it as closed and take the expected value; otherwise assume it did not close. The empirical 95% CI is bounded by the .025 and .975 quantiles of the data, summed by month.

Plot the results:

```{r, echo=F, message=F, warning=F, cache=F}
## Plot the comparison
ggplot(data=ggdat, aes(x=dt, y=act)) + geom_point(aes(color="Actual")) + geom_line(aes(color="Actual"), size=1.5, alpha=.5) + geom_point(aes(y=pred, color="JMod")) + geom_line(aes(y=pred, color="JMod")) + geom_point(aes(y=glm.pred, color="GLM")) + geom_line(aes(y=glm.pred, color="GLM")) + scale_y_continuous(name="Amount", labels=scales::dollar) + scale_x_date(name="Date", date_minor_breaks="month") + geom_ribbon(aes(ymin=`2.5%`, ymax=`97.5%`), fill="green", alpha=.1) + labs(title=paste("Solicitations under", dollar(x), "with 95% GLM CI"))
```

That's actually not bad. Of course, it would be more fair to re-fit the GLM just on FY2011-FY2015 data and use FY2016 as an out-of-sample test set, but since FY16 is still in progress it's probably not worth doing just yet.

I suspect to make big gains in predictive accuracy I'd need to use a different model structure, e.g. regressing directly on dollar amounts, which is probably worth trying later on.

# Packages used
```{r}
session_info()
```